---
output: html_document
  
---
#Final project

*[RMarkdown version](FinalProject.rmd)*
  
  
  
  ```{r,warning=F}
library(dismo)
library(maptools)
data(wrld_simpl)


```


#Set filepaths and import Basemap

```{r,warning=F}
library(maptools)
library(rgdal)
library(dismo)
library(ggplot2)
library(GISTools)
library(ggpubr) #For arranging ggplots
library(plyr)
library(dplyr)
# Set the path for outputs
output_path<-("output")
# if this folder doesn't exist, create it
if(!dir.exists(output_path)){
  dir.create(output_path)
}


# Create the folders (directories) "data" and "Final project" - If they exist already, this 
# command won't over-write them.
data_path<-(file.path("data","FinalProject"))
if(!dir.exists(data_path)){
  dir.create(data_path,recursive = TRUE)
}

#Download shapefile outline of Ghana
if(! file.exists(file.path(data_path,'countries.zip'))){
 download.file("http://www.naturalearthdata.com/http//www.naturalearthdata.com/download/10m/cultural/ne_10m_admin_0_countries.zip", dest=file.path(data_path,"countries.zip"), mode="wb") 
}
#unzip (file.path(data_path,"countries.zip"), exdir = data_path)
# Read in the shapefile with rgdal package
world <- readOGR(file.path(data_path,"ne_10m_admin_0_countries.shp"))
GhanaOutline<-(world[world$ADMIN=="Ghana",1]) #1=overall country outline




GhanaMeterGrid<-CRS("+proj=tmerc +lat_0=4.666666666666667 +lon_0=-1 +k=0.99975 +x_0=274319.51 +y_0=0 +ellps=clrk80 +towgs84=-130,29,364,0,0,0,0 +units=m +no_defs") #GhanaMeter Grid EPSG:25000
GhanaOutline




```

#Import Point Data

```{r, echo=FALSE}
#Data available at https://github.com/JPReceveur/SpatialEcologyFinalProj
MUSurvey<-read.csv("GhanaMUSurvey.csv",header=T)
MUGeo <- subset(MUSurvey, !is.na(lon) & !is.na(lat)) #Remove NAs


#plot(GhanaOutline)
# restore the box around the map
#box()
# add the points
#points(MUGeo$lon, MUGeo$lat, col='orange', pch=20, cex=0.75)
# plot points again to add a border, for better visibility
#points(MUGeo$lon, MUGeo$lat, col='red', cex=0.75)
#WSGeo

wgs1984.proj <- CRS("+proj=longlat +ellps=WGS84 +datum=WGS84 +no_defs")
MUGeo.proj <- SpatialPoints(coords=MUGeo[,c("lon","lat")], proj4string=wgs1984.proj)

plot(GhanaOutline)
points(MUGeo.proj,col="blue")


MUGeo.Meter<-spTransform(MUGeo.proj, GhanaMeterGrid) #project to Ghana meter grid
Presence<-subset(MUGeo,Total.MU.Present..VNTR.==1)#Total MU Presence coded as 0 for absence and 1 for presence
#Presence$Total.MU.Present..VNTR.
Absence<-subset(MUGeo,Total.MU.Present..VNTR.==0)
names(MUGeo)
PresenceMPM<-subset(MUGeo,Total.ER.Present==1)#Total MPM Presence coded as 0 for absence and 1 for presence
#Presence$Total.MU.Present..VNTR.
AbsenceMPM<-subset(MUGeo,Total.ER.Present==0)


#Project the subsets into WGS84 and Ghana Meter Grid

wgs1984.proj <- CRS("+proj=longlat +ellps=WGS84 +datum=WGS84 +no_defs")
Presence.proj <- SpatialPoints(coords=Presence[,c("lon","lat")], proj4string=wgs1984.proj)
Absence.proj <- SpatialPoints(coords=Absence[,c("lon","lat")], proj4string=wgs1984.proj)
PresenceMPM.proj <- SpatialPoints(coords=PresenceMPM[,c("lon","lat")], proj4string=wgs1984.proj)
AbsenceMPM.proj <- SpatialPoints(coords=AbsenceMPM[,c("lon","lat")], proj4string=wgs1984.proj)


Presence.Meter<-spTransform(Presence.proj, GhanaMeterGrid)
Absence.Meter<-spTransform(Absence.proj, GhanaMeterGrid)

PresenceMPM.Meter<-spTransform(PresenceMPM.proj, GhanaMeterGrid)
AbsenceMPM.Meter<-spTransform(AbsenceMPM.proj, GhanaMeterGrid)


#MUGeo.Meter
GhanaOutline.Meter<-spTransform(GhanaOutline,GhanaMeterGrid)

# par(mfrow=c(1,2), mai=c(0.1,0.1,0.5,0.1))
# plot(GhanaOutline,main="WGS84")
# box()
# points(MUGeo.proj,col="blue")
# 
# 
# 
# plot(GhanaOutline.Meter,main="GhanaMeterGrid")
# box()
# points(MUGeo.Meter,col="blue")

```


#Import land cover data (Ghana 2013)
```{r}

#Download land cover data from https://www.sciencebase.gov/catalog/item/5deffc05e4b02caea0f4f3fc

if(! file.exists(file.path(data_path,'LULC.zip'))){
 download.file("http://edcintl.cr.usgs.gov/downloads/sciweb1/shared/wafrica/downloads/data/west_africa_land-use_land-cover_2013_2km.zip", dest=file.path(data_path,"LULC.zip"), mode="wb") 
}
#unzip (file.path(data_path,"LULC.zip"), exdir = data_path)

LandCover2k <- stack(file.path(data_path,"west_africa_land-use_land-cover_2013_2km/swa_2013lulc_2km.tif"))
#plot(LandCover2k)
LandCover2k.Meter<-projectRaster(LandCover2k,crs=GhanaMeterGrid)
#summary(GhanaOutline.Meter)
par(mfrow=c(1,1))
r<- raster(GhanaOutline.Meter)
res(r)<-2000   #Not really sure what number to put here (units of outline shapefile are m, so just picked the resoultion for the LandCover2k.Meter file)
r[] <- rnorm(ncell(r))
#plot(r)
#plot(GhanaOutline.Meter,add=T)
r[] <- 0

GhanaRaster<-rasterize(GhanaOutline.Meter,r,getCover=T)
GhanaRaster[GhanaRaster>0]<-1
GhanaRaster[GhanaRaster==0]<-NA
#plot(GhanaRaster)
#summary(LandCover2k)

#res(LandCover2k.Meter)
LULCRaster<-LandCover2k.Meter*GhanaRaster #Subset land use by the Ghana outline (all other raster points =0)

plot(LULCRaster)
points(MUGeo.Meter,col="blue")


```

#Land use graph MU (Figure 2a)

```{r}
#pal_nlcd Function modified from https://space-lab-msu.github.io/MSUGradSpatialEcology/lab3_patch_design.html
#Table of values avialable at https://github.com/JPReceveur/SpatialEcologyFinalProj

pal_nlcd <- function() {
  data.frame(
    class = c("forest", "forest",
              "forest", "shrubland",
              "anthropogenic", "anthropogenic",
              "wetland","anthropogenic",
              "forest", "No Data",
              "Sand", "shrubland",
              "herbaceous", "herbaceous",
              "No Data", "anthropogenic",
              "shrubland", "forest",
              "anthropogenic", "anthropogenic",
              "No Data","shrubland",
              "shrubland","herbaceous",
              "Oasis","shrubland",
              "shrubland","water",
              "forest","anthropogenic",
              "No Data"),
    code = as.character(c(7, 15,
             28, 2, 8, 78,
             3,
             24, 1, 99,
             10,12, 31,
             4, 0, 27, 32,
             25, 13,
             6, 98,11,29,23,5,16,22,9,21,14,255)),
    description = c("Mangrove", "Gallery/ Riparian Forest",
                    "Swamp Forest", "Savanna", 
                    "Agriculture", "Open Mine",
                    "Wetland","Agriculture/ flood recessional",
                    "Forest", "Cloud",
                    "Sandy","Bare soil", 
                    "Herbaceous savanna", "Steppe",
                    "No data","Cropland/ oil palms", 
                    "Shrubland",  "Woodland", 
                    "Settlements","Plantation",
                    "Cloud shadow", "Rocky land",
                    "Sahelian short grass savanna", "Thicket",
                    "Oasis","Shrub and savanna",
                    "Bowe", "Water bodies",
                    "Degraded forest","Irrigated agriculture",
                    "No Data"),
    color = c("#33cccc", "#aa5ce8",
              "#beffa6", "#8cb08c", 
              "#ffff96", "#505050",
              "#000081","#ebc961", 
              "#8400a8", "#ffffff",
              "#ff99cc", "#a87000",
              "#0a9696", "#ffd09b",
              "#ffffff", "#77AD93",
              "#749373", "#28734b",
              "#ff0000", "#808000",
              "#ffffff", "#969696",
              "#a7c38d", "#f8a37b",
              "#99C147", "#77AD93",
              "#DCD939", "#3366ff",
              "#d296e6", "#cdff66",
              "##ffffff"
              ),
    stringsAsFactors = FALSE)
}


NLCD <- as.matrix(table(raster::values(LULCRaster)))
cols <- dplyr::filter(pal_nlcd(), code %in% row.names(NLCD))
par(xpd = FALSE,mai = c(0.5, 0.5, 0.2, 0)) 

plot(LULCRaster, legend=FALSE,axes=TRUE,box=FALSE,col=cols$color,
     xlab="", ylab="") 
par(xpd = TRUE) # reset to limit plotting to figure region
plot(GhanaOutline.Meter,add=T)
legend(x = -2000, y = 690000, legend = cols$description, fill = cols$color, 
       ncol = 2,  cex = 0.85, inset = 0.9) 
legend(x=500183,y=615115,legend =c("MU +","MU -"),pch=c(17,16),col=c('red','blue'))
scalebar(100000,xy = c(439429.8,61762), type="bar", divs = 2, label = c(0,50,100),below="km")
north.arrow(x=500183, y=345827, len = 10000, cex.lab = 0.75, col = "black", fg = "red")
plot(Presence.Meter,col="red",bg="red",add=T,pch=24)
plot(Absence.Meter,col="blue",bg="blue",add=T,pch=21)

# 
# dev.off()
# tiff("output/LUGhana.tiff", width = 174, height = 174, units = 'mm', res = 600)
# par(xpd = FALSE,mai = c(0.5, 0.5, 0.2, 0))
# 
# plot(LULCRaster, legend=FALSE,axes=TRUE,box=FALSE,col=cols$color,
#      xlab="", ylab="")
# par(xpd = TRUE) # reset to limit plotting to figure region
# plot(GhanaOutline.Meter,add=T)
# legend(x = -2000, y = 690000, legend = cols$description, fill = cols$color,
#        ncol = 2,  cex = 0.85, inset = 0.9)
# legend(x=450183,y=645115,legend =c("MU +","MU -"),pch=c(17,16),col=c('red','blue'))
# scalebar(100000,xy = c(439429.8,61762), type="bar", divs = 2, label = c(0,50,100),below="km")
# north.arrow(x=500183, y=345827, len = 10000, cex.lab = 0.75, col = "black", fg = "red")
# plot(Presence.Meter,col="red",bg="red",add=T,pch=24)
# plot(Absence.Meter,col="blue",bg="blue",add=T,pch=21)
# 
# dev.off()

```


#MPM Land Cover graph (Figure S1)

```{r}

#Plotted as above but showing MPM pres/abs
par(xpd = FALSE,mai = c(0.5, 0.5, 0.2, 0))

plot(LULCRaster, legend=FALSE,axes=TRUE,box=FALSE,col=cols$color,
     xlab="", ylab="")
par(xpd = TRUE) # reset to limit plotting to figure region
plot(GhanaOutline.Meter,add=T)
legend(x = -2000, y = 690000, legend = cols$description, fill = cols$color,
       ncol = 2,  cex = 0.85, inset = 0.9)
legend(x=450183,y=645115,legend =c("MPM +","MPM -"),pch=c(17,16),col=c('red4','blue'))
scalebar(100000,xy = c(439429.8,61762), type="bar", divs = 2, label = c(0,50,100),below="km")
north.arrow(x=500183, y=345827, len = 10000, cex.lab = 0.75, col = "black", fg = "red")
plot(PresenceMPM.Meter,col="red4",bg="red4",add=T,pch=24)
plot(AbsenceMPM.Meter,col="blue",bg="blue",add=T,pch=21)



# dev.off()
# tiff("output/LUGhana2.tiff", width = 174, height = 174, units = 'mm', res = 1200)
# par(xpd = FALSE,mai = c(0.5, 0.5, 0.2, 0))
# 
# plot(LULCRaster, legend=FALSE,axes=TRUE,box=FALSE,col=cols$color,
#      xlab="", ylab="")
# par(xpd = TRUE) # reset to limit plotting to figure region
# plot(GhanaOutline.Meter,add=T)
# legend(x = -2000, y = 690000, legend = cols$description, fill = cols$color,
#        ncol = 2,  cex = 0.85, inset = 0.9)
# legend(x=450183,y=645115,legend =c("MPM +","MPM -"),pch=c(17,16),col=c('red4','blue'))
# scalebar(100000,xy = c(439429.8,61762), type="bar", divs = 2, label = c(0,50,100),below="km")
# north.arrow(x=500183, y=345827, len = 10000, cex.lab = 0.75, col = "black", fg = "red")
# plot(PresenceMPM.Meter,col="red4",bg="red4",add=T,pch=24)
# plot(AbsenceMPM.Meter,col="blue",bg="blue",add=T,pch=21)
# 
# dev.off()


```

#Landscape Metrics

##Calculate 5k landscape metrics

```{r}
#Load landscape pachages
library(landscapemetrics)
library(landscapetools)
#Data same as above
MUSurvey<-read.csv("GhanaMUSurvey.csv",header=T)

MUGeo <- subset(MUSurvey, !is.na(lon) & !is.na(lat))


wgs1984.proj <- CRS("+proj=longlat +ellps=WGS84 +datum=WGS84 +no_defs")
MUGeo.proj <- SpatialPoints(coords=MUGeo[,c("lon","lat")], proj4string=wgs1984.proj)

LandCover<-LandCover2k
MUGeo.proj<-spTransform(MUGeo.proj,crs(LandCover))
GhanaOutline2<-spTransform(GhanaOutline,crs(LandCover))



#Sample 5k buffer (5000 m) around each sample point using landscape metrics
Buffer5k<-sample_lsm(
  LandCover,
  MUGeo.proj,
  plot_id = NULL,
  shape = "square",
  size=5000,
  all_classes = FALSE,
  return_raster = FALSE,
  verbose = TRUE,
  progress = FALSE,level="landscape",type="aggregation metric")

Buffer5k
#unique(Buffer$metric)
```


##Landscape Metric MU 5k

```{r}
#For loop testing differences between PresAbs values for the chosen metrics (MetricList)
MetricList<-c("contag","np","lsi","cohesion")

for(i in 1:length(MetricList)){
  Subset<-subset(Buffer5k,Buffer5k$metric==MetricList[i])
  print(unique(Subset$metric))
  
  
  
  Subset$plot_id<-MUGeo$USITEID
  Subset$PresAbs<-MUGeo$Total.MU.Present..VNTR.
  Subset$PresAbs[Subset$PresAbs==1]<-"Present"
  Subset$PresAbs[Subset$PresAbs==0]<-"Absent"
  Plot <- ggplot(Subset,aes( x=PresAbs,y=value,col=PresAbs))+ylab(MetricList[i])+xlab("MU VNTR Profiling")+geom_boxplot()+scale_color_manual(values=c("blue","red"))+theme(legend.position="none")
  print(Plot)
  print(kruskal.test(data=Subset,value~PresAbs))
  Trtdata <- ddply(Subset, c("metric","PresAbs"), summarise,
                 N    = length(value),
                 mean = mean(value),
                 sd   = sd(value),
                 se   = sd / sqrt(N)
  )
  print(Trtdata)
  
}

#LSI Landscape shape index #standardized measure of total edge LSI=1 =single patch increases from there pg 158 Fragstats

#NP <- number of patches pg 149
#contagion -sum of proportional abundance of each patch, inverse to edge density pg154


```

##Landscape Metric MPM 5k

```{r}

MetricList<-c("contag","np","lsi","cohesion")
#Same as above but using Total.ER.Present to calculate for MPM abundance
for(i in 1:length(MetricList)){
  Subset<-subset(Buffer5k,Buffer5k$metric==MetricList[i])
  print(unique(Subset$metric))
  
  
  
  Subset$plot_id<-MUGeo$USITEID
  Subset$PresAbs<-MUGeo$Total.ER.Present #Total ER presence coded as 1= presence, 0 =absence
  Subset$PresAbs[Subset$PresAbs==1]<-"Present" 
  Subset$PresAbs[Subset$PresAbs==0]<-"Absent"
  Plot <- ggplot(Subset,aes( x=PresAbs,y=value,col=PresAbs))+ylab(MetricList[i])+xlab("MPM VNTR Profiling")+geom_boxplot()+scale_color_manual(values=c("blue","red4"))+theme(legend.position="none")
  print(Plot)
  print(kruskal.test(data=Subset,PresAbs~value))
  
}



```

#10k landscape metrics MU

```{r}
#Data same as above
MUSurvey<-read.csv("GhanaMUSurvey.csv",header=T)

MUGeo <- subset(MUSurvey, !is.na(lon) & !is.na(lat))


#Project to WGS84
wgs1984.proj <- CRS("+proj=longlat +ellps=WGS84 +datum=WGS84 +no_defs")
MUGeo.proj <- SpatialPoints(coords=MUGeo[,c("lon","lat")], proj4string=wgs1984.proj)

LandCover<-LandCover2k
MUGeo.proj<-spTransform(MUGeo.proj,crs(LandCover))
GhanaOutline2<-spTransform(GhanaOutline,crs(LandCover))



#Calculate 10 k buffer (10000 m) around sample points for landscape metrics
Buffer10k<-sample_lsm(
  LandCover,
  MUGeo.proj,
  plot_id = NULL,
  shape = "square",
  size=10000,
  all_classes = FALSE,
  return_raster = FALSE,
  verbose = TRUE,
  progress = FALSE,level="landscape",type="aggregation metric")

#Buffer10k
#unique(Buffer$metric)
```


##Landscape Metric MU 10 k

```{r}
  
MetricList<-c("contag","np","lsi","cohesion")

for(i in 1:length(MetricList)){
  Subset<-subset(Buffer10k,Buffer10k$metric==MetricList[i])
  print(unique(Subset$metric))
  
  
  
  Subset$plot_id<-MUGeo$USITEID
  Subset$PresAbs<-MUGeo$Total.MU.Present..VNTR.
  Subset$PresAbs[Subset$PresAbs==1]<-"Present"
  Subset$PresAbs[Subset$PresAbs==0]<-"Absent"
  Plot <- ggplot(Subset,aes( x=PresAbs,y=value,col=PresAbs))+ylab(MetricList[i])+xlab("MU VNTR Profiling")+geom_boxplot()+scale_color_manual(values=c("blue","red"))+theme(legend.position="none")
  print(Plot)
  print(kruskal.test(data=Subset,value~PresAbs))
  Trtdata <- ddply(Subset, c("metric","PresAbs"), summarise,
                 N    = length(value),
                 mean = mean(value),
                 sd   = sd(value),
                 se   = sd / sqrt(N)
  )
  print(Trtdata)
  
}

#LSI Landscape shape index #standardized measure of total edge LSI=1 =single patch increases from there pg 158 Fragstats

#NP <- number of patches pg 149
#contagion -sum of proportional abundance of each patch, inverse to edge density pg154

#Subset for Contag only plot
Contag<-subset(Buffer10k,Buffer10k$metric=="contag")
Contag$plot_id<-MUGeo$USITEID
Contag$PresAbs<-MUGeo$Total.MU.Present..VNTR.
Contag$PresAbs[Contag$PresAbs==1]<-"Present"
Contag$PresAbs[Contag$PresAbs==0]<-"Absent"

Plot <- ggplot(Contag,aes( x=PresAbs,y=value,col=PresAbs,shape=PresAbs))+ylab(MetricList[i])+xlab("MU VNTR Profiling")+geom_boxplot()+scale_color_manual(values=c("blue","red"))+theme(legend.position="none")+ylab("Contagion Index")+geom_point()
Plot
  

# 
# dev.off()
# tiff("output/ContagPlot.tiff", width = 84, height = 84, units = 'mm', res = 600)
# Plot
# dev.off()



  
```

#Landscape Metric MPM Presence Absence
##5 kilometer

```{r}
#List of metrics to run kruskal.wallis tests on
MetricList<-c("contag","np","lsi","cohesion")

for(i in 1:length(MetricList)){
  Subset<-subset(Buffer5k,Buffer5k$metric==MetricList[i])
  print(unique(Subset$metric))
  
  
  
  Subset$plot_id<-MUGeo$USITEID #unique site ID from file
  Subset$PresAbs<-MUGeo$Total.ER.Present
  Subset$PresAbs[Subset$PresAbs==1]<-"Present"
  Subset$PresAbs[Subset$PresAbs==0]<-"Absent"
  Plot <- ggplot(Subset,aes( x=PresAbs,y=value,col=PresAbs))+ylab(MetricList[i])+xlab("MPM  Profiling")+geom_boxplot()+scale_color_manual(values=c("blue","red4"))+theme(legend.position="none")
  #print(Plot)
  print(kruskal.test(data=Subset,value~PresAbs))
  Trtdata <- ddply(Subset, c("metric","PresAbs"), summarise,
                 N    = length(value),
                 mean = mean(value),
                 sd   = sd(value),
                 se   = sd / sqrt(N)
  )
  print(Trtdata)
  
}



```

##10k MPM landscape metrics

```{r}

MetricList<-c("contag","np","lsi","cohesion")

for(i in 1:length(MetricList)){
  Subset<-subset(Buffer10k,Buffer5k$metric==MetricList[i])
  print(unique(Subset$metric))
  
  
  
  Subset$plot_id<-MUGeo$USITEID
  Subset$PresAbs<-MUGeo$Total.ER.Present
  Subset$PresAbs[Subset$PresAbs==1]<-"Present"
  Subset$PresAbs[Subset$PresAbs==0]<-"Absent"
  Plot <- ggplot(Subset,aes( x=PresAbs,y=value,col=PresAbs))+ylab(MetricList[i])+xlab("MPM  Profiling")+geom_boxplot()+scale_color_manual(values=c("blue","red4"))+theme(legend.position="none")
  #print(Plot)
  print(kruskal.test(data=Subset,value~PresAbs))
  Trtdata <- ddply(Subset, c("metric","PresAbs"), summarise,
                 N    = length(value),
                 mean = mean(value),
                 sd   = sd(value),
                 se   = sd / sqrt(N)
  )
  print(Trtdata)
  
}



```





#Overview Plots (Figure 1a-d)
Figures 1a-b

```{r}

Overview<-MUGeo

#Total.MU.Present..VNTR. coded as 1 =pres,0=abs
Overview$PresAbsMU[Overview$Total.MU.Present..VNTR.==1]<-"MU +"
Overview$PresAbsMU[Overview$Total.MU.Present..VNTR.==0]<-"MU -"

Overview$PresAbsMPM[Overview$Total.ER.Present==1]<-"MPM +"
Overview$PresAbsMPM[Overview$Total.ER.Present==0]<-"MPM -"
theme_set(theme_bw(base_size = 12)+theme(panel.grid.major = element_blank(), panel.grid.minor = element_blank()))

#Count of MU by region
MUCount<-ggplot(Overview,aes( x=PresAbsMU,fill=PresAbsMU))+geom_histogram(stat="count")+facet_grid(~REGION)+scale_fill_manual(values=c('blue','red'))+ylab("Count By Region")+xlab("Presence of MU at site")+theme(legend.position = "none")+ theme(axis.text.x = element_text(angle = 45, hjust = 1))+stat_count(aes(y=..count.., label=..count..), geom="text", vjust=-.5) 

#Count of MPM by region
MPMCount<-ggplot(Overview,aes( x=PresAbsMPM,fill=PresAbsMPM))+geom_histogram(stat="count")+facet_grid(~REGION)+scale_fill_manual(values=c('blue','red4'))+ylab("Count By Region")+xlab("Presence of MPM at site")+theme(legend.position = "none")+ theme(axis.text.x = element_text(angle = 45, hjust = 1))+stat_count(aes(y=..count.., label=..count..), geom="text", vjust=-.5) 

ggarrange(MUCount,MPMCount,
          labels = c("a", "b"), ncol = 1,nrow=2)


# dev.off()
# tiff("output/MU.MPMCountByRegion.tiff", width = 174, height = 174, units = 'mm', res = 1200)
# ggarrange(MUCount,MPMCount,
#           labels = c("a", "b"), ncol = 1,nrow=2)
# dev.off()

```

Figures 1c-d

```{r}
#Extract Land cover from locations where there is a sample point
Patch<-extract_lsm(LandCover, y = MUGeo.proj, what = "lsm_p_area")

unique(Patch$class)
#Transform values to LU classes, list available at https://github.com/JPReceveur/SpatialEcologyFinalProj
Patch$class[Patch$class ==8]<-"Agriculture"
Patch$class[Patch$class ==2]<-"Savanna"
Patch$class[Patch$class ==3]<-"Wetland"
Patch$class[Patch$class ==13]<-"Settlements"
Patch$class[Patch$class ==21]<-"Degraded forest"
Patch$class[Patch$class ==1]<-"Forest"
Patch$class[Patch$class ==78]<-"Open Mine"
Patch$class[Patch$class ==15]<-"Gallery/Riparian forest"

Patch$extract_id<-MUGeo$ï..USITEID
Patch$MUPresAbs<-as.character(MUGeo$Total.MU.Present..VNTR.)
Patch$MUPresAbs[Patch$MUPresAbs==0]<-"MU -"
Patch$MUPresAbs[Patch$MUPresAbs==1]<-"MU +"

Patch$MPMPresAbs<-as.character(MUGeo$Total.ER.Present)
Patch$MPMPresAbs[Patch$MPMPresAbs==0]<-"MPM -"
Patch$MPMPresAbs[Patch$MPMPresAbs==1]<-"MPM +"

#MU presence by Land use
MUByLocation<-ggplot(Patch,aes( x= MUPresAbs,fill=MUPresAbs))+geom_histogram(stat="count")+facet_grid(~class)+scale_fill_manual(values=c('blue','red'))+ylab("Count By Land Use")+xlab("Presence of MU at site")+theme(legend.position = "none")+ theme(axis.text.x = element_text(angle = 45, hjust = 1))+stat_count(aes(y=..count.., label=..count..), geom="text", vjust=-.5) 


#MPM pres/abs by land use
MPMByLocation<-ggplot(Patch,aes( x= MPMPresAbs,fill=MPMPresAbs))+geom_histogram(stat="count")+facet_grid(~class)+scale_fill_manual(values=c('blue','red4'))+ylab("Count By Land Use")+xlab("Presence of MPM at site")+theme(legend.position = "none")+ theme(axis.text.x = element_text(angle = 45, hjust = 1))+stat_count(aes(y=..count.., label=..count..), geom="text", vjust=-.5) 



ggarrange(MUByLocation,MPMByLocation,
          labels = c("a", "b"), ncol = 1,nrow=2)

# 
# dev.off()
# tiff("output/PresenceByLU.tiff", width = 174, height = 174, units = 'mm', res = 600)
# ggarrange(MUByLocation,MPMByLocation,
#           labels = c("c", "d"), ncol = 1,nrow=2)
# 
# 
# 
# dev.off()



```







#Import Climate Data

```{r}
#Climate Data downloaded 
#if(! file.exists(file.path(data_path,'ClimateData.zip'))){
# download.file("https://biogeo.ucdavis.edu/data/worldclim/v2.1/base/wc2.1_30s_bio.zip", dest=file.path(data_path,"ClimateData.zip"), mode="wb") 
#}
#unzip (file.path(data_path,"ClimateData.zip"), exdir = data_path)
#Folder Climate Data subset contains
#wc2.1_30s_bio_1.tif Annual Mean Temperature
#wc2.1_30s_bio_12.tif Annual Precipitation
#wc2.1_30s_bio_15.tif Precipitation seasonality
#wc2.1_30s_bio_7.tif Temperature Annual range
#wc2.1_30s_bio_14.tif Precipitation of driest month

ClimateTif<-list.files(file.path(data_path,"ClimateDataSubset"), pattern='tif$', full.names=TRUE )
ClimateSubset <- stack(ClimateTif)


ma.area <- extent(GhanaOutline)

ClimateSubsetGhana <- crop(ClimateSubset, ma.area)
GhanaClimateSubset.Meter<-projectRaster(ClimateSubsetGhana,crs=GhanaMeterGrid)


```







#SDM Models

```{r}
#Import survey data and seperate into pres absence data
MUSurvey<-read.csv("GhanaMUSurvey.csv",header=T)
MUGeo <- subset(MUSurvey, !is.na(lon) & !is.na(lat))
Presence<-subset(MUGeo,Total.MU.Present..VNTR.==1)
Presence$Total.MU.Present..VNTR.
Absence<-subset(MUGeo,Total.MU.Present..VNTR.==0)

wgs1984.proj <- CRS("+proj=longlat +ellps=WGS84 +datum=WGS84 +no_defs")
Presence.proj <- SpatialPoints(coords=Presence[,c("lon","lat")], proj4string=wgs1984.proj)
Absence.proj <- SpatialPoints(coords=Absence[,c("lon","lat")], proj4string=wgs1984.proj)


Presence.Meter<-spTransform(Presence.proj, GhanaMeterGrid)
Absence.Meter<-spTransform(Absence.proj, GhanaMeterGrid)


#MUGeo.Meter


presvals <- extract(GhanaClimateSubset.Meter, Presence.Meter)
absvals <- extract(GhanaClimateSubset.Meter, Absence.Meter)
pb <- c(rep(1, nrow(presvals)), rep(0, nrow(absvals)))
sdmdata <- data.frame(cbind(pb, rbind(presvals, absvals)))
#head(sdmdata)

#pairs(sdmdata[,2:5], cex=0.1)



PresLat<-Presence.Meter$lat
PresLong<-Presence.Meter$lon

PresLongLat<-data.frame(PresLong,PresLat)


pred_nf<-GhanaClimateSubset.Meter

#See dismo vignette 
set.seed(0)
group <- kfold(PresLongLat, 5)
pres_train <- PresLongLat[group != 1, ]
pres_test <- PresLongLat[group == 1, ]

ext<-extent(GhanaOutline.Meter)

AbsLat<-Absence.Meter$lat
AbsLong<-Absence.Meter$lon

AbsLongLat<-data.frame(AbsLong,AbsLat)


pred_nf<-GhanaClimateSubset.Meter


set.seed(0)
group <- kfold(AbsLongLat, 5)
#group
abs_train <- AbsLongLat[group != 1, ]

abs_test <- AbsLongLat[group == 1, ]


r <- raster(pred_nf, 1)
# plot(!is.na(r), col=c('white', 'light grey'), legend=FALSE)
# plot(ext, add=TRUE, col='red', lwd=2)
# points(abs_train, pch='-', cex=0.5, col='black')
# points(abs_test, pch='-',  cex=0.5, col='black')
# points(pres_train, pch= '+', col='green')
# points(pres_test, pch='+', col='blue')
```


##Bioclim model MU pres/abs

```{r}
bc<-bioclim(pred_nf,pres_train) #pres_train

#plot(bc, a=1,b=2,p=0.85)
```

```{r}
e<-evaluate(pres_test,abs_test,bc,pred_nf)
e

tr<-threshold(e,'spec_sens')
#tr

pb<-predict(pred_nf,bc,ext=ext,progress='')
pb

par(mfrow=c(1,2))
plot(pb, main='Bioclim, raw values')
plot(GhanaOutline.Meter, add=TRUE, border='dark grey')
plot(pb > tr, main='presence/absence')
plot(GhanaOutline.Meter, add=TRUE, border='dark grey')
points(pres_train, pch='+')

```

##Domain Model
```{r}
dm <- domain(pred_nf, pres_train)
e <- evaluate(pres_test, abs_test, dm, pred_nf)
e

pd = predict(pred_nf, dm, ext=ext, progress='')
par(mfrow=c(1,2))
plot(pd, main='Domain, raw values')
plot(GhanaOutline.Meter, add=TRUE, border='dark grey')
tr <- threshold(e, 'spec_sens')
plot(pd > tr, main='presence/absence')
plot(GhanaOutline.Meter, add=TRUE, border='dark grey')
points(pres_train, pch='+')

```





```{r}
predictors<-GhanaClimateSubset.Meter
colnames(pres_train)<-colnames(abs_train)
train <- rbind(pres_train, abs_train)


pb_train <- c(rep(1, nrow(pres_train)), rep(0, nrow(abs_train)))
envtrain <- extract(predictors, train)
envtrain <- data.frame( cbind(pa=pb_train, envtrain) )
#head(envtrain)

testpres <- data.frame( extract(predictors, pres_test) )
testabs <- data.frame( extract(predictors, abs_test) )
```

##GLM Model MU pres/abs

```{r}
# logistic regression:
#names(envtrain)
gm1 <- glm(pa ~ wc2.1_30s_bio_1+wc2.1_30s_bio_12+wc2.1_30s_bio_15+wc2.1_30s_bio_7+wc2.1_30s_bio_14, family = binomial(link = "logit"), data=envtrain)

summary(gm1)
#coef(gm1)
gm2 <- glm(pa ~ wc2.1_30s_bio_1+wc2.1_30s_bio_12+wc2.1_30s_bio_15+wc2.1_30s_bio_7+wc2.1_30s_bio_14,
           family = gaussian(link = "identity"), data=envtrain)
evaluate(testpres,testabs,gm1)

ge2 <- evaluate(testpres, testabs, gm1)
ge2

pg <- predict(predictors, gm1, ext=ext)

#plot(pg, main='GLM/binomial, raw values')
#plot(GhanaOutline.Meter, add=TRUE, border='dark grey')
tr <- threshold(ge2, 'spec_sens')

plot(pg > tr, main='presence/absence MU (GLM)')
plot(GhanaOutline.Meter, add=TRUE, border='dark grey')

legend(x=500183,y=615115,legend =c("MU +","MU -"),pch=c(17,16),col=c('red','blue'))
scalebar(100000,xy = c(439429.8,61762), type="bar", divs = 2, label = c(0,50,100),below="km")
north.arrow(x=500183, y=345827, len = 10000, cex.lab = 0.75, col = "black", fg = "red")
points(pres_test,col="red",bg="red",pch=24)
points(abs_test,col="blue",bg="blue",pch=21)

# 
# 
# dev.off()
# tiff("output/GLMSDMWTestSet.tiff", width = 174, height = 174, units = 'mm', res = 1200)
# plot(pg > tr, main='presence/absence MU (GLM)')
# plot(GhanaOutline.Meter, add=TRUE, border='dark grey')
# legend(x=500183,y=615115,legend =c("MU +","MU -"),pch=c(17,16),col=c('red','blue'))
# scalebar(100000,xy = c(439429.8,61762), type="bar", divs = 2, label = c(0,50,100),below="km")
# north.arrow(x=500183, y=345827, len = 10000, cex.lab = 0.75, col = "black", fg = "red")
# points(pres_test,col="red",bg="red",pch=24)
# points(abs_test,col="blue",bg="blue",pch=21)
# dev.off()



```


##Random Forest SDM Model MU

```{r}
library(randomForest)
model <- pa ~ wc2.1_30s_bio_1+wc2.1_30s_bio_12+wc2.1_30s_bio_15+wc2.1_30s_bio_7+wc2.1_30s_bio_14
rf1 <- randomForest(model, data=envtrain)
model <- factor(pa) ~ wc2.1_30s_bio_1+wc2.1_30s_bio_12+wc2.1_30s_bio_15+wc2.1_30s_bio_7+wc2.1_30s_bio_14
rf2 <- randomForest(model, data=envtrain)
erf <- evaluate(testpres, testabs, rf1)
erf

pr <- predict(predictors, rf1, ext=ext)
plot(pr, main='Random Forest, regression')
plot(GhanaOutline.Meter, add=TRUE, border='dark grey')
tr <- threshold(erf, 'spec_sens')
plot(pr > tr, main='presence/absence MU (RF)')
plot(GhanaOutline.Meter, add=TRUE, border='dark grey')
legend(x=500183,y=615115,legend =c("MU +","MU -"),pch=c(17,16),col=c('red','blue'))
scalebar(100000,xy = c(439429.8,61762), type="bar", divs = 2, label = c(0,50,100),below="km")
north.arrow(x=500183, y=345827, len = 10000, cex.lab = 0.75, col = "black", fg = "red")
points(pres_test,col="red",bg="red",pch=24)
points(abs_test,col="blue",bg="blue",pch=21)

importance(rf2)


# dev.off()
# tiff("output/rfSDMWTestSet.tiff", width = 174, height = 174, units = 'mm', res = 1200)
# plot(pr > tr, main='presence/absence MU (RF)')
# plot(GhanaOutline.Meter, add=TRUE, border='dark grey')
# legend(x=500183,y=615115,legend =c("MU +","MU -"),pch=c(17,16),col=c('red','blue'))
# scalebar(100000,xy = c(439429.8,61762), type="bar", divs = 2, label = c(0,50,100),below="km")
# north.arrow(x=500183, y=345827, len = 10000, cex.lab = 0.75, col = "black", fg = "red")
# points(pres_test,col="red",bg="red",pch=24)
# points(abs_test,col="blue",bg="blue",pch=21)
# 
# dev.off()

```

##Support Vector machines

```{r}

library(kernlab)
svm <- ksvm(pa ~ wc2.1_30s_bio_1+wc2.1_30s_bio_12+wc2.1_30s_bio_15+wc2.1_30s_bio_7+wc2.1_30s_bio_14, data=envtrain)

esv <- evaluate(testpres, testabs, svm)
esv

ps <- predict(predictors, svm, ext=ext)
par(mfrow=c(1,2))
plot(ps, main='Support Vector Machine')
plot(GhanaOutline.Meter, add=TRUE, border='dark grey')
tr <- threshold(esv, 'spec_sens')
plot(ps > tr, main='presence/absence')
plot(GhanaOutline.Meter, add=TRUE, border='dark grey')
points(pres_train, pch='+')
points(abs_train, pch='-', cex=0.25)

```

##Combining model predictions (Final two models (rf,glm) shown seperately and not combined)
```{r}
# Final 
# models <- stack(pg, pr )
# names(models) <- c("glm", "rf")
# plot(models)
# 
# m <- mean(models)
# plot(m, main='average score')
# 
# auc <- sapply(list(ge2, erf), function(x) x@auc)
# w <- (auc-0.5)^2
# m2 <- weighted.mean( models[[c("glm", "rf")]], w)
# plot(m2)
# plot(GhanaOutline.Meter,add=T)
# plot(Presence.Meter,col="red",bg="red",add=T,pch=24)
# plot(Absence.Meter,col="blue",bg="blue",add=T,pch=21)
```





#Correlograms MU/MPM presence absence by distance (Figure 2b)

```{r}
library(ncf)
#Not used Moran I assumption data is normal
# 
# qqnorm(MUGeo$Total.ER.Present)
# qqline(MUGeo$Total.ER.Present,col="blue")
# 
# 

cor.MUPos <- correlog(MUGeo.proj$lat, MUGeo.proj$lon, MUGeo$Total.MU.Present..VNTR., increment = 10000, resamp = 100)
plot(cor.MUPos,xlab= "Distance (m, mean of class)",ylab="Correlation (MU positivity)",main="")
abline(h = 0, col = "purple")

# 
# dev.off()
# tiff("output/CorrelogramMUPresence.tiff", width = 84, height = 84, units = 'mm', res = 1200)
# plot(cor.MUPos,xlab= "Distance (m, mean of class)",ylab="Correlation (MU positivity)",main="")
# abline(h = 0, col = "purple")
# 
# dev.off()


cor.MPMPos <- correlog(MUGeo.proj$lat, MUGeo.proj$lon, MUGeo$Total.ER.Present, increment = 10000, resamp = 100)

plot(cor.MPMPos,xlab= "Distance (m, mean of class)",ylab="Correlation (MPM positivity)",main="")
abline(h = 0, col = "purple")
# 
# dev.off()
# tiff("output/CorrelogramMPMPresence.tiff", width = 84, height = 84, units = 'mm', res = 1200)
# plot(cor.MPMPos,xlab= "Distance (m, mean of class)",ylab="Correlation (MPM positivity)",main="")
# abline(h = 0, col = "purple")
# 
# dev.off()

```


